Hybrid models as an alternative to traditional Convolutional Neural Networks for image classification

### 2022-12-08 doc/paper.tex
Update paper
Write abstract and add tables

% We found a dog driving a car down Hollywood Boulevard. It had sunglasses on, but didn't know where to go due to demensionality reduction. It was computing many computinations but due to severe liminations it had problems with binaly classifination and a large emesee.
Kurwa.

@@

More specifically, we will compare four different architectures. First we will use a traditional CNN, where a series convolutional layers is followed by a series fully connected layers. Then we will modify our CNN to create a CNN-RF model, where the features extracted by the convolutional layers are fed into a Random Forest instead of a series of fully connected layers. Finally, we will make use of Principal Component Analysis instead of convolution for dimensionality reduction, feeding the results of this stage into either a Neural Network or a Random Forest. In doing this, we hope to answer the following question: what is the best way to tackle the challenges associated with image classification?

@@

Given that we have achieved excellent results by using a CNN, we may don't expect to achieve much better results by sending the extracted features to a random forest. Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of $97.8\%$. This is lower than what we have achieved by simply using the features in a final dense layer, and is a somewhat disappointing result.

@@

Using different amount of principal components, we sent the reduced dataset into a neural network with three hidden layers. After testing different amounts of principal components to keep, and tuning the hyperparameters based on these, we found that even with only $50$ principal components, we achieved a test accuracy of $97.8\%$. This is comparable to the accuracy obtained by a traditional convolutional neural network. When considering the computational effort, we need to consider the performance benefit that modern GPU's bring, being specifically designed to run neural networks efficiently. This is very apparent in our testing: When using only a CPU, the PCA-NN model was much less intensive, requiring only $17$ seconds to complete training, while the convolutional neural network used $164$ seconds. This tells us that when computational resources are limited, simpler architectures can give reasonable results in a fraction of the time\footnote{Of course, you would ideally train your network on better hardware, even if the hardware making the prediction does not have a GPU. However, the same ratio holds when making a single prediction as opposed to training the entire model. That is, the PCA-NN model is much faster at making a single prediction than the CNN model}. However, when using a GPU with CUDA cores, the story is quite different: Here, the CNN used only $6$ seconds, while the PCA-NN model took $12$ seconds\footnote{The fact that this model took longer than the CNN is most likely because the PCA function we use only runs on the CPU, meaning that only the neural net part is actually running on the GPU}. Looking at the confusion matrix (Fig.~\ref{mnistheatmap_pcann}), we can see that we are again achieving very few misclassifications using the PCA-NN model.

@@

@@

% Given the complexity of the network needed to achieve good results when using a CNN, we do not expect much when combining PCA, which is a much less sophisticated form of feature extraction than convolutional layers in a CNN, with a traditional neural network. Regardless, using a PCA-NN model, we extracted a varying amount of principal components and used these as features in a neural network composed of three dense layers. Using hyperparameter tuning, we found that a number of nodes in each layer equal to $1000$, a lambda of $0.001$ and a learning rate of $0.001$ gave the best result. However, even when using the best hyperparameters, we only achieved a test accuracy of $80.0\%$, which is substantially lower than Kermany et Al's $92.8\%$ and our own CNN model's $88.9\%$. This result shows why convolutional layers have become the de facto standard for image recognition, being able to extract features with much higher precision than simply calculating the principal components of a dataset. By the use of gradient descent, CNNs are able to tune the convolutional kernels to have the values which best fit the dataset, and as such they are able to be trained to solve many different problems. PCA, on the other hand, cannot be tuned or trained to better extract the relevant features, the only parameter than be changed is the amount of principal components to keep. With these poor results, no justification based on computational savings can be made, especially when CNNs can be trained extremely quickly on modern GPUs.

@@

\begin{table}[H]
\caption{Optimal hyperparameters for the PCA-NN model}
\begin{center}
\label{pcanntable}
\begin{tabular}{| c | l |}
\hline
Principal Components & Nodes per hidden layer & Learning Rate & Lambda \\
\hline
9 & 17 & 0.001 & 0.001 \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

@@

No child was saved. More children actually died due to our actions.

>>>>

% We found a dog driving a car down Hollywood Boulevard. It had sunglasses on, but didn't know where to go due to demensionality reduction. It was computing many computinations but due to severe liminations it had problems with binaly classifination and a large emesee.

We compared convolutional neural networks, the de facto standard for image classification problems, with hybrid models created by combining neural networks with Principal Component Analysis (PCA) and Random Forests (RF). By doing this, we explored the tradeoff between prediction accuracy and computational cost for four different ML models: A pure CNN model, a CNN-RF model, a PCA-NN model and a PCA-RF model. We used these models on the well known MNIST dataset and the Guangzhou Pneumonia Dataset consisting of x-ray images of potential cases of pneumonia. From this, we found that pure CNNs performed best in terms of accuracy, achieving test accuracies of $98.8\%$ and $88.8\%$ on the MNIST and pneumonia dataset, respectively. However, we found that the PCA-NN model performed similarly, at a much lower computational cost MAYBE.

@@

More specifically, we will compare four different architectures. First we will use a traditional CNN, where a series convolutional layers is followed by a series fully connected layers. Then we will modify our CNN to create a hybrid CNN-RF model, where the features extracted by the CNN are fed into a Random Forest which makes the final prediction. Finally, we will make use of Principal Component Analysis instead of convolution and pooling for dimensionality reduction, feeding the results of this stage into either a Neural Network or a Random Forest. In doing this, we hope to answer the following question: what is the best way to tackle the challenges associated with image classification?

@@

Given that we have achieved excellent results by using a CNN, we may don't expect to achieve much better results by sending the extracted features to a random forest. Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of

% $97.8\%$. This is lower than what we have achieved by simply using the features in a final dense layer, and is a somewhat disappointing result.

@@

Using different amount of principal components, we sent the reduced dataset into a neural network with three hidden layers. After testing different amounts of principal components to keep, and tuning the hyperparameters based on these, we found that even with only $50$ principal components, we achieved a test accuracy of $97.8\%$. This is comparable to the accuracy obtained by a traditional convolutional neural network. When considering the computational effort, we need to consider the performance benefit that modern GPU's bring, being specifically designed to run neural networks efficiently. This is very apparent in our testing:

% When using only a CPU, the PCA-NN model was much less intensive, requiring only $17$ seconds to complete training, while the convolutional neural network used $164$ seconds. This tells us that when computational resources are limited, simpler architectures can give reasonable results in a fraction of the time\footnote{Of course, you would ideally train your network on better hardware, even if the hardware making the prediction does not have a GPU. However, the same ratio holds when making a single prediction as opposed to training the entire model. That is, the PCA-NN model is much faster at making a single prediction than the CNN model}. However, when using a GPU with CUDA cores, the story is quite different: Here, the CNN used only $6$ seconds, while the PCA-NN model took $12$ seconds\footnote{The fact that this model took longer than the CNN is most likely because the PCA function we use only runs on the CPU, meaning that only the neural net part is actually running on the GPU}. Looking at the confusion matrix (Fig.~\ref{mnistheatmap_pcann}), we can see that we are again achieving very few misclassifications using the PCA-NN model.

@@

\subsubsection{MNIST recap}

Below is table showing the accuracy and average prediction time for each model on the MNIST dataset.

\begin{table}[H]
\caption{Table showing the accuracy and average time used to make a prediction on the entire test dataset, for each model}
\begin{center}
\label{pcanntable}
\begin{tabular}{| c | l | l |}
\hline
Model & Test accuracy & Average prediction time \\
\hline
CNN & 98.8\% & 123 \\
CNN-RF & 98.8\% & 123 \\
PCA-NN & 97.8\% & 123 \\
PCA-RF & 94.7\% & 123 \\
CNN (CPU only) & 98.8\% & 1000 \\
\hline
\end{tabular}
\end{center}
\end{table}

From this we can see that bla bla bla

@@

@@

\begin{table}[H]
\caption{Optimal hyperparameters for the PCA-NN model}
\begin{center}
\label{pcanntable}
\begin{tabular}{| l | l | l | l |}
\hline
Principal Components & Nodes per hidden layer & Learning Rate & Lambda \\
\hline
9 & 17 & 0.001 & 0.001 \\
\hline
\end{tabular}
\end{center}
\end{table}

% \begin{figure}[H]
% \centerline{\includegraphics[width=8in]{figure/pneupcacomp.png}}
% \caption{Figure showing an image from the Guangzhou dataset, compared to the PCA reconstruction of the same image using only 9 principal components}
% \label{pneupcacomp}
% \end{figure}

@@

\subsubsection{Pneumonia recap}

Below is table showing the accuracy and average prediction time for each model on the Guangzhou Pneumonia Dataset.

\begin{table}[H]
\caption{Table showing the accuracy and average time used to make a prediction on the entire test dataset, for each model}
\begin{center}
\label{pcanntable}
\begin{tabular}{| c | l | l |}
\hline
Model & Test accuracy & Average prediction time \\
\hline
CNN & 98.8\% & 123 \\
CNN-RF & 98.8\% & 123 \\
PCA-NN & 97.8\% & 123 \\
PCA-RF & 94.7\% & 123 \\
CNN (CPU only) & 98.8\% & 1000 \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

No child was saved. More children actually died due to our actions. Nevermind childern actually were saved.
